---
title: "다변량통계분석 hw1"
author: "2019150418 통계학과 김나연"
date: '2021 3 19 '
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Q1
```{r}
X=data.frame(X1=c(4,12,8,10,6,8),X2=c(2,3,0,2,1,4),X3=c(5,9,6,8,6,2))
attach(X)
```

##(a)
```{r}
xbar=colMeans(X) ; xbar=as.matrix(xbar)
S=cov(X)
xbar
S
```
##(b)
```{r}
R=cor(X)
R
```

##(c)
```{r}
Z1=-X1+3*X2-2*X3
c(mean(Z1),var(Z1))
```

##(d)
```{r}
Z2=5*X2-X3
c(mean(Z2),var(Z2))
```

##(e)
```{r}
Z3=-X1+X3
c(mean(Z3),var(Z3))
```

##(f)
```{r}
Zprime=data.frame(Z1,Z2,Z3)
Zprime=as.matrix(Zprime)
zt=t(Zprime)
cov(zt)
```

# Q2
```{r}
hw1=read.table('C:/Users/user/Desktop/고려대/3-1/다변량통계분석/usair.dat',header=T)
attach(hw1)
p=ncol(hw1)
```

## (a)
```{r}
xbar=colMeans(hw1); xbar=as.matrix(xbar)
S=cov(hw1)
xbar
S
```
## (b)
```{r}
R=cor(hw1)
R
```
## (c)
```{r}
for (i in 1:p){
  qqnorm(hw1[,i],xlab='Quantiles of standard normal',ylab=colnames(hw1)[i])
  qqline(hw1[,i],col='pink')
}
apply(hw1,2,shapiro.test)
```
### By seeing qqplots, the variable WIND and DAYS are following univariate normal distribution. So we should transform the other variables. 


## (d)
```{r}
dis2=mahalanobis(hw1,xbar,S)
dis2
```
## (e)
```{r}
library(lattice)
qqmath(dis2,distribution=function(p) qchisq(p,df=7),
       panel=function(x,...) {panel.qqmathline(x,...)
       panel.qqmath(x,...)})
shapiro.test(dis2)
```
### By seeing plot, there are too much outliers. So we can judge this data does not follow a multivariate normal distribution, 
## (f)
```{r}
library(forecast)
library(dplyr)
hw1_f=hw1 %>% select(c(SO2,TEMP,MANUF,POP,PRECIP))
hw1_f
lambda=rep(0,5)
for (i in 1:ncol(hw1_f)){
     lambda[i]=BoxCox.lambda(hw1_f[,i],method='loglik')
}
lambda
```
### ABOUT THE VARIABLE SO2
```{r}
bc1=BoxCox(hw1_f[,1],lambda[1])
qqnorm(bc1)
qqline(bc1)
shapiro.test(bc1)
```
### ABOUT THE VARIABLE TEMP
```{r}
bc2=BoxCox(hw1_f[,2],lambda[2])
qqnorm(bc2)
qqline(bc2)
shapiro.test(bc2)
```

### ABOUT THE VARIABLE MANUF
```{r}
bc3=BoxCox(hw1_f[,3],lambda[3])
qqnorm(bc3)
qqline(bc3)
shapiro.test(bc3)
```

### ABOUT THE VARIABLE POP
```{r}
bc4=BoxCox(hw1_f[,4],lambda[4])
qqnorm(bc4)
qqline(bc4)
shapiro.test(bc4)
```

### ABOUT THE VARIABLE PRECIP
```{r}
bc5=BoxCox(hw1_f[,5],lambda[5])
qqnorm(bc5)
qqline(bc5)
shapiro.test(bc5)
```
#### By seeing the plots and the result of the shapiro test after doing an appropriate transformation, we can know that each variable follows a univariate normal distribution.
